---
title: "Analysis predictivo proyecto 1"
output: html_notebook
---

```{r}
library("dplyr")
library("ggplot2")
library("vcd")
library("tidyverse")
library("gplots")
library("reshape2")
library("ade4") # ACM
library("factoextra") # fancy ACM
```

```{r}
data <- read.csv("diabetes_012_health_indicators_BRFSS2015.csv")
```

Visualize the dataset
```{r}
head(data)
```

# Cleaning of the dataset #

```{r}

# Calculate the number of person in each type of diabete
freq_table <- table(data$Diabetes_012)

# Create a barplot
barplot(freq_table, col="blue", xlab = "Diabete types", ylab="Number of people", main = "Number of people for each type of diabete")

#Create a pie chart
label_percent <- paste(round(100 * freq_table / sum(freq_table), 1), "%")
pie(freq_table, col = c("blue", "red", "green"), labels = label_percent, main = "Distribution of Diabetes Types", cex.main = 1.2)
legend("topright", legend = c("No Diabetes", "Prediabetes", "Diabetes"), fill = c("blue", "red", "green"),
       title = "Diabetes Types", cex = 0.8)


freq_table
```


We choose to put together prediabete and diabete, as we consider that in both case people are sick.
```{r}
data$Diabetes_012[data$Diabetes_012 == 2]=1

#we calculate again the number of person in each type of diabete
freq_table <- table(data$Diabetes_012)

# Create a barplot
barplot(freq_table, col="blue", xlab = "Diabete", ylab="Number of people", main = "Number of people for each type of diabete")

#Create a pie chart
label_percent <- paste(round(100 * freq_table / sum(freq_table), 1), "%")
pie(freq_table, col = c("blue", "red"), labels = label_percent, main = "Distribution of Diabetes Types", cex.main = 1.2)
legend("topright", legend = c("No Diabetes", "Diabetes"), fill = c("blue", "red"),
       title = "Diabetes Types", cex = 0.8)

freq_table
```

We change the name from the column Diabetes_012 to diabete, and we turn it in factors
```{r}
data$diabete <- data$Diabetes_012
data <- data[-which(names(data) == "Diabetes_012")] 

data$diabete = as.factor(data$diabete)
```




We then look for missing values
```{r}
# Count the number of NA in every column
na_counts <- colSums(is.na(data))

print(na_counts)
```

# Exploratory analysis #

## Categorical variables ##

We take a look at the importance of the categorical variables on diabete: we take a look at which variables have a lot of 1 when people have diabete, and which variable are important when people don't have diabete.

We start with the sick ones.
```{r}
# We create a new dataframe with only categorical variables for people with diabete
temp_1= data %>% filter(diabete==1)
temp_1=temp_1 %>% select(-BMI, -GenHlth, -MentHlth, -PhysHlth, -Age, -Education, -Income, -diabete)

# Reshape the data for ggplot
data_long <- reshape2::melt(temp_1)

# Create a graph with ggplot
ggplot(data_long, aes(x = value, fill = variable)) +
  geom_bar(position = "dodge") +
  labs(title = "Frequency of 0 and 1 for categorical variables for sick people",
       x = "Values (0 or 1)",
       y = "Frequency") +
  theme_minimal()

```

We now do the same for people who don't have diabete.
```{r}
# We create a new dataframe with only categorical variables for people without diabete
temp_0= data %>% filter(diabete==0)
temp_0=temp_0 %>% select(-BMI, -GenHlth, -MentHlth, -PhysHlth, -Age, -Education, -Income, -diabete)

# Reshape the data for ggplot
data_long <- reshape2::melt(temp_0)

# Create a graph with ggplot
ggplot(data_long, aes(x = value, fill = variable)) +
  geom_bar(position = "dodge") +
  labs(title = "Frequency of 0 and 1 for categorical variables for healthy people",
       x = "Values",
       y = "Frequency") +
  theme_minimal()

```

We now want to compare for each categorical variable the percentage of sick people compare to the percentage of healthy people.
```{r}
#Create a datframe with categorical variables for people with and withou diabete
cat=data %>% select(-BMI, -MentHlth, -PhysHlth, -Age, -Education, -Income)
nb_col = ncol(cat)

#Create barplots or each variable
for (i in 1:(nb_col - 1)) { #-1 to avoid the column diabete
  colonne <- names(cat)[i]
  
  p =ggplot(cat, aes_string(x = colonne, fill = "diabete")) +
    geom_bar(position = "dodge") +
    labs(title = paste("Frequency of 0 and 1 for", colonne),
         x = "Values",
         y = "Frequency") +
    scale_fill_manual(values = c("1" = "blue", "0" = "red")) +
    theme_minimal()
  
  p <- p + geom_text(stat = "count", aes(label = scales::percent(..prop..), y = ..prop..), position = position_dodge(width = 0.9), vjust = -0.5, color = "white")
  
  print(p)
}
```

## Numerical variables ##


```{r}
# Histogram for BMI
ggplot(data, aes(x=BMI)) +
  geom_histogram(alpha=0.5) +
  labs(title = "Histogram of the BMI variable",
         x = "BMI",
         y = "Counts") +
  NULL

#Histogram for MentHlth
ggplot(data, aes(x=MentHlth)) +
  geom_histogram(alpha=0.5) +
  labs(title = "Histogram of the MentHlth variable",
         x = "MentHlth",
         y = "Counts") +
  NULL

#Histogram for PhysHlth
ggplot(data, aes(x=PhysHlth)) +
  geom_histogram(alpha=0.5) +
  labs(title = "Histogram of the PhysHlth variable",
         x = "PhysHlth",
         y = "Counts") +
  NULL
```


## Further exploration ##


```{r}
# List of variable names to exclude in the models : the binaries variables
variables_excluded <- c("HighBP", "HighChol", "CholCheck", "Smoker", "Stroke", "HeartDiseaseorAttack", "PhysActivity", "Fruits", "Veggies", "HvyAlcoholConsump", "AnyHealthcare", "NoDocbcCost", "DiffWalk", "Sex")

# Loop through the column names of the dataset
# If the variable is not in the list of variables of interest, move to the next one

for (i in setdiff(names(data), c("diabete", variables_excluded))) {

  # Create an ordered version of the current categorical variable
  data[[i]] <- ordered(data[[i]])
  
  # Perform logistic regression to assess the trend effect and display the summary of the logistic regression
  model <- glm(diabete ~ as.numeric(data[[i]]), data = data, family = binomial(link = "logit"))
  cat(paste("Variable:", i, "\n"))
  summary(model)
  
  # Create a trend plot
  plot(data[[i]], model$fitted.values, xlab = i, ylab = "Probability of Diabetes", main = paste("Trend of", i))
}
```

# Correlation analysis #

We transorm the categorical variables in factors
```{r}
names_bin = data %>% select_if(
  function(x) is.numeric(x) & length(unique(x)) == 2) %>% names()

for (col_name in names_bin) {
  data[[col_name]] <- as.factor(data[[col_name]])
}
```

```{r}
names_cat = c("GenHlth","Age","Education","Income")
for (col_name in names_cat) {
  data[[col_name]] <- as.factor(data[[col_name]])
}
```

```{r}
names_num = c("BMI","MentHlth","PhysHlth")
```


We use some functions to calculate the Cramer's V coefficient
```{r}
cramer_v = function(df, x1, x2) {
  tab = table(df[[x1]], df[[x2]])
  res = as.matrix(tab) %>% rstatix::cramer_v()
  return(res)
}
```

```{r}
cramer_cors = function(df, col_names) {
  df_cor = combn(col_names, 2) %>% 
    t() %>% 
    as.data.frame() %>% 
    set_names(c("Var1", "Var2"))
  df_cor = df_cor %>% mutate(
    cramer_v = map2_dbl(Var1, Var2, function(x,y) cramer_v(df, x, y))
  ) %>% arrange(-cramer_v)
  return(df_cor)
}
```

```{r} 
names_bin <- c(names_bin, "diabete")
```

```{r}
res = cramer_cors(data, c(names_bin, names_cat))
res
```

We create the Cramer's V correlation matrix.
```{r}
variables_uniques <- unique(c(res$Var1, res$Var2))
matrice_correlation <- matrix(0, nrow = length(variables_uniques), ncol = length(variables_uniques))
rownames(matrice_correlation) <- colnames(matrice_correlation) <- variables_uniques

for (i in 1:nrow(res)) {
  var1 <- res$Var1[i]
  var2 <- res$Var2[i]
  correlation <- res$cramer_v[i]
  matrice_correlation[var1, var2] <- correlation
  matrice_correlation[var2, var1] <- correlation  
}

diag(matrice_correlation) <- 1

matrice_correlation_long <- melt(matrice_correlation)
```

Matrix display
```{r}
ggplot(matrice_correlation_long, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "yellow") +
  labs(title = "Cramer's V correlation matrix") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

We select the results with our target variable "diabete".
```{r}
resultats_filtres <- res %>%
  filter(Var1 == "diabete" | Var2 == "diabete")

print(resultats_filtres)
```

We calculate the correlation for our numerical variables with the Spearman method.
```{r}
a=data[,c("BMI","MentHlth","PhysHlth")]
a[["BMI"]]=as.numeric(a[["BMI"]])
a[["MentHlth"]]=as.numeric(a[["MentHlth"]])
a[["PhysHlth"]]=as.numeric(a[["PhysHlth"]])

cor_mat_spearman <- cor(a, method = "spearman")

print(cor_mat_spearman)
```
```{r}
gg <- GGally::ggcorr(
  a, method=c("pairwise","spearman"),  
  label=T, hjust=1, label_size=2, layout.exp=10, size=3)

gg + ggtitle("Correlation matrix with numerical variables")
```

## MCA ##

We use a MCA to confirm our results
```{r}
# Selecting specific columns from the dataframe (binary/categorical variables)
data_binaries <- data[, c("diabete", "HighBP", "HighChol", "CholCheck", "Smoker", "Stroke", "HeartDiseaseorAttack", "PhysActivity", "Fruits", "Veggies", "HvyAlcoholConsump", "AnyHealthcare", "NoDocbcCost", "DiffWalk", "Sex")]

# Performing Multiple Correspondence Analysis (MCA) on the selected binary/categorical variables
acm <- dudi.acm(data_binaries, scannf = FALSE, nf = 4)
```

```{r}
# Creating a scree plot to visualize the eigenvalues
screeplot(acm, main = "Screeplot of Multiple Correspondence Analysis of categorical variables")
```

```{r}
# Displaying a summary of the Multiple Correspondence Analysis
summary(acm)
# Calculating and displaying the inertia of the MCA
inertia.dudi(acm)
```

```{r}
# Creating a scree plot with eigenvalues using 'fviz_screeplot' function
fviz_screeplot(acm, choice = "eigenvalue")
```

```{r}
# Visualizing generally the contribution of variables to the dimensions of MCA
fviz_mca_var(acm, repel = TRUE)
```

```{r}
# Visualizing the contributions of variables to each axis (dimension) individually on barplots
fviz_contrib(acm, choice = "var", axes = 1)
fviz_contrib(acm, choice = "var", axes = 2)
fviz_contrib(acm, choice = "var", axes = 3)
fviz_contrib(acm, choice = "var", axes = 4)
```